---
title: '<center> <h1>Linear regression model in R using TensorFlow</h1></center>'
author: '<center> <h3>Vadim I. Nazarov</h3> </center> <center> <h4><vdm.nazarov@gmail.com></h4> </center>'
output: 
  html_document: 
    theme: spacelab
---

#### Preface
This is an R version of the wonderful TensorFlow example made by [Aymeric Damien](https://github.com/aymericdamien) in [this project](https://github.com/aymericdamien/TensorFlow-Examples/).

Original TensorFlow example notebook: [link](https://github.com/aymericdamien/TensorFlow-Examples/blob/master/notebooks/2_BasicModels/linear_regression.ipynb)

TensorFlow R examples website: [link](https://vadimnazarov.github.io/tensorflow-r-examples/)

TensorFlow R examples GitHub: [link](https://github.com/vadimnazarov/tensorflow-r-examples/)

#### Installation
Installation instructions for TensorFlow R package: [link](https://rstudio.github.io/tensorflow/)

#### Libraries
```{r libs}
library(tensorflow)
library(ggplot2)
```

#### Parameters
```{r params}
set.seed(20)
learning_rate = 0.1
training_epochs = 1000
verbose_step = 50
```

#### Training data
```{r data}
x_data = c(3.3,4.4,5.5,6.71,6.93,4.168,9.779,6.182,7.59,2.167,
           7.042,10.791,5.313,7.997,5.654,9.27,3.1)
y_data = c(1.7,2.76,2.09,3.19,1.694,1.573,3.366,2.596,2.53,1.221,
           2.827,3.465,1.65,2.904,2.42,2.94,1.3)
n_samples = length(x_data)
```

#### TensorFlow graph input
```{r input}
x = tf$placeholder(tf$float32)
y = tf$placeholder(tf$float32)
```

#### Initial weights for the model
```{r weights}
W = tf$Variable(tf$random_uniform(shape(1L), -1.0, 1.0), name = "weight")
b = tf$Variable(tf$zeros(shape(1L)), name = "bias")
```

#### Linear model description
```{r model}
lr_model = tf$mul(W, x) + b
```

#### Loss function - mean squared error
```{r loss}
loss_fun = tf$reduce_sum((lr_model - y) ^ 2) / (2 * n_samples)
```

#### Optimisation using the gradient descent algorithm
```{r optim}
optim_fun = tf$train$GradientDescentOptimizer(learning_rate)$minimize(loss_fun)
```

#### Initialise all variables
```{r init}
init = tf$initialize_all_variables()
```

#### Launch the graph and plot the result
```{r run, fig.align="center"}
with(tf$Session() %as% sess, {
  sess$run(init)
  
  #### For every epoch
  for (epoch in 1:training_epochs) {
    #### Fit all training data
    for (i in 1:n_samples) {
      sess$run(optim_fun, feed_dict = dict(x = x_data[i], y = y_data[i]))
    }
    
    if ((epoch + 1) %% verbose_step == 0) {
      loss_val = sess$run(loss_fun, feed_dict = dict(x = x_data, y = y_data))
      cat(paste0("Epoch: ", format(epoch+1, width = 4), 
                 " loss=", format(loss_val, width = 10), 
                 " W=", format(sess$run(W), width = 8), 
                 " b=", format(sess$run(b), 8), "\n"))
    }
  }
  
  training_loss = sess$run(loss_fun, feed_dict = dict(y = x_data, x = y_data))
  cat(paste0("Resulting loss=", format(training_loss, width = 10), 
             " W=", format(sess$run(W), width = 8), 
             " b=", format(sess$run(b), 8), "\n"))
  
  df = data.frame(X = x_data, Y = y_data)
  
  x_start = min(x_data)
  x_end   = max(x_data)
  y_start = sess$run(W) * x_start + sess$run(b)
  y_end   = sess$run(W) * x_end + sess$run(b)
  
  ggplot() + 
    geom_point(aes(x = X, y = Y), data = df) + 
    geom_segment(aes(x = x_start, 
                 xend = x_end,
                 y = y_start,
                 yend = y_end)) +
  theme_bw()
})
```