---
title: '<center> <h1>Nearest neighbor model in R using TensorFlow</h1></center>'
author: '<center> <h3>Vadim I. Nazarov</h3> </center>'
output: 
  html_document: 
    theme: spacelab
---

#### Preface
This is an R version of the wonderful TensorFlow example made by [Aymeric Damien](https://github.com/aymericdamien) in [this project](https://github.com/aymericdamien/TensorFlow-Examples/).

Original TensorFlow example notebook: [link](https://github.com/aymericdamien/TensorFlow-Examples/blob/master/notebooks/2_BasicModels/nearest_neighbor.ipynb)

Code for this document: [link](https://github.com/vadimnazarov/tensorflow-r-examples/src/2_BasicModels/nearest_neighbor.Rmd)

TensorFlow R examples website: [link](https://vadimnazarov.github.io/tensorflow-r-examples/)

TensorFlow R examples GitHub: [link](https://github.com/vadimnazarov/tensorflow-r-examples/)

#### Installation
Installation instructions for TensorFlow R package: [link](https://rstudio.github.io/tensorflow/)

#### Libraries
```{r libs}
library(tensorflow)
```

#### Parameters
```{r params}
train_size = 5000
test_size = 200
n_features = 784L
```

#### Training data
```{r data}
input_data = tf$contrib$learn$datasets$mnist
mnist = input_data$read_data_sets("./tmp/data", one_hot = T)

train_data = mnist$train$next_batch(train_size)
x_train = train_data[[1]]
y_train = train_data[[2]]

test_data = mnist$test$next_batch(test_size)
x_test = test_data[[1]]
y_test = test_data[[2]]
```

#### TensorFlow graph input
```{r input}
x_src = tf$placeholder(tf$float32, shape(NULL, n_features))
x_tgt = tf$placeholder(tf$float32, shape(n_features))
```

#### Nearest neighbors calculation using L1 distance
```{r weights}
distance = tf$reduce_sum(tf$abs(x_src - x_tgt), reduction_indices = 1L)
# L2 distance will look like this
# distance = tf$reduce_sum((x_src - x_tgt) ^ 2, reduction_indices = 1L)
```

#### Nearest neighbors model description
```{r model}
nn_model = tf$arg_min(distance, 0L)
```

#### Initialise all variables
```{r init}
init = tf$initialize_all_variables()
acc = .0
```

#### Launch the L1 distance graph and print the resulting accuracy
```{r run}
with(tf$Session() %as% sess, {
  sess$run(init)
  
  # For every object
  for (i in 1:nrow(x_test)) {
    # Get the nearest object's class
    nn_index = sess$run(nn_model, feed_dict = dict(x_src = x_train, x_tgt = x_test[i,])) + 1
    pred = which.max(y_train[nn_index, ]) - 1
    real = which.max(y_test[i, ]) - 1
    if (pred == real) {
      acc = acc + 1 / nrow(x_test)
    } else {
      cat("test no. ", i, " pred: ", pred, " true: ", real, "\n")
    }
  }
  
  cat("Accuracy: ", acc, "\n")
})
```