---
title: '<center> <h1>Basic operations in R TensorFlow</h1></center>'
author: '<center> <h3>Vadim I. Nazarov</h3> </center> '
output: 
  html_document: 
    theme: spacelab
---

#### Preface
This is an R version of the wonderful TensorFlow example made by [Aymeric Damien](https://github.com/aymericdamien) in [this project](https://github.com/aymericdamien/TensorFlow-Examples/).

Original TensorFlow example notebook: [link](https://github.com/aymericdamien/TensorFlow-Examples/blob/master/notebooks/1_Introduction/basic_operations.ipynb)

Code for this document: [link](https://github.com/vadimnazarov/tensorflow-r-examples/src/1_Introduction/basic_operations.Rmd)

TensorFlow R examples website: [link](https://vadimnazarov.github.io/tensorflow-r-examples/)

TensorFlow R examples GitHub: [link](https://github.com/vadimnazarov/tensorflow-r-examples/)

#### Installation
Installation instructions for TensorFlow R package: [link](https://rstudio.github.io/tensorflow/)

#### Libraries
```{r libs}
library(tensorflow)
```

#### Operations with constants
```{r const}
# The value returned by the constructor represents 
# the output of the Constant operation.
a = tf$constant(2)
b = tf$constant(3)
```

#### Launch (i.e., compute / execute) the default graph
```{r defgraph}
# "with" allows us to initialize (and later remove) the session "sess" only for the code
# in the inner code block and don't ruin any external variables.
with(tf$Session() %as% sess, {
  cat("a = 2, b = 3\n")
  cat("Addition with constants: ", sess$run(a + b), "\n")
  cat("Multiplication with constants: ", sess$run(a * b), "\n")
})
```

#### Operations with variables - graph inputs
```{r variables}
# The value returned by the constructor represents 
# the output of the Variable operation. This is the "boxes"
# for the input data.
a = tf$placeholder(tf$int16)
b = tf$placeholder(tf$int16)

# Now define what we will do with this variables.
add_fun_long = tf$add(a, b)
# The TensorFlow R library provides nice straightforward aliases 
# for TensorFlow operations like addition, multiplication, etc.
# More info here: https://rstudio.github.io/tensorflow/using_tensorflow_api.html
add_fun_short = a + b
mul_fun_long = tf$mul(a, b)
mul_fun_short = a * b
```

#### Launch the default graph
```{r defgraph2}
with(tf$Session() %as% sess, {
  cat("Addition with variables #1: ", sess$run(add_fun_long, feed_dict = dict(a = 3, b = 5)), "\n")
  cat("Addition with variables #2: ", sess$run(add_fun_short, feed_dict = dict(a = 3, b = 5)), "\n")
  cat("Multiplication with variables #1: ", sess$run(mul_fun_long, feed_dict = dict(a = 3, b = 5)), "\n")
  cat("Multiplication with variables #2: ", sess$run(mul_fun_short, feed_dict = dict(a = 3, b = 5)), "\n")
})
```

#### Matrix multiplication
```{r matrix}
# Matrix Multiplication from TensorFlow official tutorial

# 1x2 matrix
matrix1 = tf$constant(matrix(c(3, 3), nrow = 1))
# 2x1 matrix
matrix2 = tf$constant(matrix(c(2, 2), nrow = 2))

# MATrix MULtiplication operation
matmul = tf$matmul(matrix1, matrix2)
```

#### Launch the default graph
```{r defgraph3}
with(tf$Session() %as% sess, {
  print(sess$run(matmul))
})
```