---
title: '<center> <h1>"Hello, World!" in R TensorFlow</h1></center>'
author: '<center> <h3>Vadim I. Nazarov</h3> </center> '
output: 
  html_document: 
    theme: spacelab
---

#### Preface
This is an R version of the wonderful TensorFlow example made by [Aymeric Damien](https://github.com/aymericdamien) in [this project](https://github.com/aymericdamien/TensorFlow-Examples/).

Original TensorFlow example notebook: [link](https://github.com/aymericdamien/TensorFlow-Examples/blob/master/notebooks/1_Introduction/helloworld.ipynb)

Code for this document: [link](https://github.com/vadimnazarov/tensorflow-r-examples/src/1_Introduction/helloworld.Rmd)

TensorFlow R examples website: [link](https://vadimnazarov.github.io/tensorflow-r-examples/)

TensorFlow R examples GitHub: [link](https://github.com/vadimnazarov/tensorflow-r-examples/)

#### Installation
Installation instructions for TensorFlow R package: [link](https://rstudio.github.io/tensorflow/)

#### Libraries
```{r libs}
library(tensorflow)
```

#### Create new Constant with the string
```{r const}
# Create the Constant op. with the helloworld string.
# The op. will be added to the default graph.
hello = tf$constant("Hello, TensorFlow!")
```

#### Start the session
```{r sess}
# In the future we will find the more convenient way to start sessions.
sess = tf$Session()
```

#### Launch the default graph
```{r defgraph}
sess$run(hello)
```